Graph Neural Networks For Multi-Image Matching

### 2019-10-03 latex/iclr2020/main.tex
Final Latex for ICLR - forgot to commit earlier (as usual)

\begin{abstract}
    In geometric computer vision applications, multi-image feature matching gives more accurate and robust solutions compared to simple two-image matching.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence can be difficult or expensive to acquire.
    Geometric consistency losses are added to aid training, though unlike optimization based methods no geometric information is necessary at inference time.
    To the best of our knowledge, no other works have used graph neural networks for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D feature positions from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \citep{lowe2004distinctive} followed by RANSAC \citep{fischler1981random} have been the standard for decades.
However RANSAC-based approaches are limited to matching pairs of images, which can lead to global inconsistencies in the matching.
% TODO: Mention trifocal tensors
Other works, such as \cite{wang2017multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent.

However, these multi-view consistency algorithms struggle in distributed and robust settings.
Having image features suited for this task would help improve performance, and deep learning has revolutionized how image features are computed \citep{yi2016lift}.
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.

@@

In this work, we address these concerns using Graph Neural Networks (GNNs).
The proposed method works directly on the graph of correspondences between the image features, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss, the cycle consistency loss, to train the network, thus avoiding the difficulty of expensive hand-labeling.
We use geometric consistency losses to aid training, though no geometric information is used at inference time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss \citep{zhou2015multi, leonardos2016distributed}.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster to achieve comparable accuracy than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K dataset \citep{li2010location} to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets such as \cite{zagoruyko2015learning}; \cite{yi2016lift}; and \cite{brachmann2017dsac}.
This approach is limited if one does not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as \cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, \cite{yi2018learning} attempts to improve correspondences by learning match probabilities for two-view RANSAC for greater robustness and speed.
Like us \cite{zhu2017unpaired} use cycle consistency in their loss; however their method is for image generation and is restricted to pairwise cycle consistency.
Our method can be applied to 3 or more images. 
\cite{hartmann2017learned} in fact learns multi-image matching but requires heavy supervision from 3D object reconstructions, which are not available in the context we are in.

\subsection{Graph Neural Networks}
Graph neural networks, true to their name, are deep neural networks operating in graph domains.
Multi-image matching is novel application of them, as they are more typically used for applications such as citation networks or recommendation systems.
They have received more attention recently \citep{bronstein2017geometric, defferrard2016convolutional, kipf2017semi, scarselli2009graph, gama2018mimo, gama2018convolutional, battaglia2018relational}.
The first methods to learn neural networks over graphs were the so-called Spectral methods.
They used the eigenvectors of the graph Laplacian to compute convolutions, which can learn graph specific convolution kernels (as in \cite{bruna2013spectral}), but require an a-priori known graph structure. 
Newer non-spectral methods do not require a-priori knowledge, as seen in \cite{bronstein2017geometric, kipf2017semi, scarselli2009graph}; and \cite{gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
\cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in \cite{battaglia2018relational}.
The closest work on Graph Neural Networks to ours is \cite{kipf2016variational}, which uses GNNs to reconstruct the adjacency matrices, though not in a geometric context.

@@

Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
An outline of our approach can be seen in figure \ref{fig:pipeline}.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Graphs $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ have a set of vertices $\mathcal{V}$ and of edges $\mathcal{E} \subseteq \mathcal{V} \times \mathcal{V}$ and implicitly in the subsequent discussions we assume the edges are directed.
For a vertex $v \in \mathcal{V}$ we use $\mathcal{N}^{h}(v)$ to denote the $h$-hop neighbors of $v$, with the superscript left out for 1-hop neighbors.
Similarly $\mathcal{E}(v)$ is used to denote the edges associated with $v$.
To denote the vertices connected to an edge $e \in \mathcal{E}$ we write $e(v_1, v_2)$.
\subsection{Correspondence Graph} \label{sec:corrgraph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, with an associated adjacency matrix $\mA$.
The graph is constructed from putative correspondences of image features across images, typically constructed using feature descriptor distance (e.g. SIFT feature distance).
While there are many interesting methods for computing these putative correspondences \citep{suh2015subgraph, yi2018learning}, we do not explore them in this work.
Typically putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

@@

\subsection{Graph Neural Networks} \label{sec:gnns}
As input to our method we are given a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with the features described in Section \ref{sec:corrgraph}: $\vf_v \; \forall v \in \mathcal{V}$ and $\vf_e \; \forall e \in \mathcal{E}$.
As with any neural network, GNNs have layered outputs.
We describe the output of layer $k$ as $\vf_v^{(k)} \in \sR^{m_k} \, \forall v \in \mathcal{V}$ and $\vf_e^{(k)} \in \sR^{p_k}\, \forall e \in \mathcal{E}$, with the initial embeddings denoted $\vf_v^{(0)} = \vf_v$ and $\vf_e^{(0)} = \vf_e$.
To aid future analysis, we will represent the features as matrices, denoting the vertex embedding matrix as $\mF_V^{(k)}$ and the edge embedding matrix as $\mF_E^{(k)}$.
If a superscript is not specified then it refers to the final output of the network.

First we describe older methods of GNNs to give context, then we describe the method we use in this work.
Many older methods assume we have the adjacency matrix $\mA$ of the graph known a-priori \cite{bruna2013spectral}, and can encode graph convolutions using the eigenvectors of $\mA$ (these are known as spectral methods).
However, we do not have this luxury, as the correspondence structure changes from image set to image set, and thus we use non-spectral Graph Neural Networks.
Newer models use non-spectral methods, which often ultimately amount to transforming each node with learned weights then averaging each node's representation with its neighbors, known as a message pass \citep{kipf2017semi, defferrard2016convolutional, gama2018mimo, gama2018convolutional}.
Some works such as \cite{gama2019convolutional} use pooling operations on the vertices to make the graph smaller and thus aid computation, but as we need labels on every vertex of the original graph, we cannot use this.
Most GNNs used in these works can be expressed mathematically as:
\begin{align*}
\tilde{\vf}_v^{(k+1)} =&\; \sigma\left( b^{(k)} + \mW_0^{k} \vf_{v}^{(k)} + \sum_{h=0}^H \sum_{v' \in \mathcal{N}^{h}(v)} f_{e(v,v')} \mW_{h}^{k} \vf_{v'}^{(k)} \right) \\
\end{align*}
The weights/biases $\mW_{h}^{k}$, $b^{(k)}$ are all learned, with no learning done on the edge weights $f_{e(v,v')}$. 
Note that this is just averages over $h$-hop neighborhoods, where the weights on the edges remain static through the computation.
Given that we are trying to prune edges, it would be sensible to add features over edges to learn which ones to prune and which to keep such as in \cite{scarselli2009graph}.

In this work we use the method and implementation described in \cite{battaglia2018relational} (more specifically repetitions of the architecture described in \cite{battaglia2016interaction}).
\cite{battaglia2018relational} uses message passes between node features as well as edge features, which the model can use to prune unnecessary or erroneous edges.
Therefore there is intermediate processing on the edges before information is passed to the vertices.

@@

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth correspondences but using unsupervised methods and geometric side losses.
Thus we test against ground truth correspondence graph adjacency matrices computed from the bundle adjustment output.
Our method performs better than 25 iteration of the MatchALS \citep{zhou2015multi} method, but does not perform as well as 50 iterations.
We do not perform as well as the Projected Gradient Descent Doubly Stochastic (PGDDS) \citep{leonardos2016distributed} but we perform significantly faster than them.
We perform better than a simple eigenvalue based method \citep{pachauri2013solving}.
Note that we perform much better in $L_1$ performance rather than $L_2$, as we optimized the network weights using an $L_1$ loss.
}
\label{tab:rome16ktab}
\end{table*}

@@

For these experiments we train with the $L_1$ norm and geometric consistency losses.
We evaluate on a test set using the ground truth adjacency matrix, which we compute from the bundle adjustment given by the Rome16K dataset.
However, we do not train with the ground truth adjacency matrix, only with the noisy adjacency matrix from the graph.
We also add the geometric loss (\ref{eq:geom_cost2}) which helps improve testing performance (see figure \ref{fig:geomloss}).
We use the $L_1$ and ROC AUC metrics to measure performance.
For this method to work, we need the dimension of the embedding to be at least the number of unique points in the scene.
Picking the correct number is difficult a-priori, and is a problem with all cycle consistency based methods.
Here we use the ground truth dimension of the embedding to test both our method and the baselines.

@@

\subsubsection{Graffiti Dataset} \label{sec:graffiti}
We run our trained model on the more Graffiti Dataset (formatted to be able to be input to our model properly).
The Graffiti Dataset is the most common benchmark used in feature matching algorithms (e.g. \ref{leonardos2016distributed, zhou2015multi}).
The results are shown in figure \ref{fig:errorlines} in the rightmost figure.
As the graffiti dataset is very small (only 6 views total), we were not able to train on it.
We randomly permute the intra-image order of the features to add some variance - by design the GNN outputs the same result each time, while the optimization methods have a very small amount of variance.
The transferred results of Graffiti are similar to the test error of Rome16K - smaller $L_1$ error and comparable ROC error.
This shows that the GNNs trained in Rome16K generalize similarly to the optimization based methods.

>>>>

\begin{abstract}
    In geometric computer vision applications, multi-image feature matching gives more accurate and robust solutions compared to simple two-image matching.
    In this work, we formulate multi-image matching as a graph embedding problem, then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence can be difficult or expensive to acquire.
    Geometric consistency losses are added to aid training, though unlike optimization based methods no geometric information is necessary at inference time.
    To the best of our knowledge, no other works have used graph neural networks for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D feature positions from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \citep{lowe2004distinctive} combined with RANSAC \citep{fischler1981random} have been the standard for decades.
However RANSAC-based approaches are limited to matching pairs of images, which can lead to global inconsistencies in the matching.
% TODO: Mention trifocal tensors
Other works, such as \cite{wang2017multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent.

However, these multi-view consistency algorithms struggle in distributed and noisy settings.
Having image features suited for this task would help improve performance, and deep learning has revolutionized how image features are computed \citep{yi2016lift}.
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.

@@

In this work, we address these concerns using Graph Neural Networks (GNNs).
The proposed method works directly on the graph of correspondences between the image features, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss, the cycle consistency loss, to train the network, thus avoiding the difficulty of expensive hand-labeling.
We use geometric consistency losses to aid training, though no geometric information is used at inference time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss \citep{zhou2015multi, leonardos2016distributed}.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster to achieve comparable accuracy than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K dataset \citep{li2010location} to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised multi-view cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets such as \cite{zagoruyko2015learning}; \cite{yi2016lift}; and \cite{brachmann2017dsac}.
This approach is limited if one does not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as \cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, \cite{yi2018learning} attempts to improve correspondences by learning match probabilities for two-view RANSAC for greater robustness and speed.
Like us \cite{zhu2017unpaired} use cycle consistency in their loss; however their method is for image generation and is restricted to pairwise cycle consistency.
Our method can be applied to 3 or more images. 
\cite{hartmann2017learned} learns multi-image matching but requires heavy supervision from 3D object reconstructions, which can be difficult or expensive to obtain.

\subsection{Graph Neural Networks}
Graph neural networks, true to their name, are deep neural networks operating in graph domains.
Multi-image matching is novel application of them, as they are more typically used for applications such as citation networks or recommendation systems.
They have received more attention recently \citep{bronstein2017geometric, defferrard2016convolutional, kipf2017semi, scarselli2009graph, gama2018mimo, gama2018convolutional, battaglia2018relational}.
The first methods to learn neural networks over graphs were the so-called Spectral methods.
They used the eigenvectors of the graph Laplacian to compute convolutions, which can learn graph specific convolution kernels (as in \cite{bruna2013spectral}), but require an a-priori known graph structure. 
Newer non-spectral methods do not require a-priori knowledge, as seen in \cite{bronstein2017geometric, kipf2017semi, scarselli2009graph}; and \cite{gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
\cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in \cite{battaglia2018relational}.
The closest work on Graph Neural Networks to ours is \cite{kipf2016variational}, which uses GNNs to reconstruct adjacency matrices, though not in a geometric context.

@@

Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
An outline of our approach can be seen in figure \ref{fig:pipeline}.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Graphs $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ have a set of vertices $\mathcal{V}$ and of directed edges $\mathcal{E} \subseteq \mathcal{V} \times \mathcal{V}$.
For a vertex $v \in \mathcal{V}$ we use $\mathcal{N}^{h}(v)$ to denote the $h$-hop neighbors of $v$, with the superscript left out for 1-hop neighbors.
Similarly $\mathcal{E}(v)$ is used to denote the edges associated with $v$.
To denote the vertices connected to an edge $e \in \mathcal{E}$ we write $e(v_1, v_2)$.
\subsection{Correspondence Graph} \label{sec:corrgraph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, with an associated adjacency matrix $\mA$.
The graph is constructed from putative correspondences of image features across images, typically constructed using feature descriptor distance (e.g. SIFT feature distance).
While there are many interesting methods for computing these putative correspondences \citep{suh2015subgraph, yi2018learning}, we do not explore them in this work.
Typically putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

@@

\subsection{Graph Neural Networks for Feature Matching} \label{sec:gnns}
As input to our method we are given a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with the features described in Section \ref{sec:corrgraph}: $\vf_v \; \forall v \in \mathcal{V}$ and $\vf_e \; \forall e \in \mathcal{E}$.
As with any neural network, GNNs have layered outputs.
We describe the output of layer $k$ as $\vf_v^{(k)} \in \sR^{m_k} \, \forall v \in \mathcal{V}$ and $\vf_e^{(k)} \in \sR^{p_k}\, \forall e \in \mathcal{E}$, with the initial embeddings denoted $\vf_v^{(0)} = \vf_v$ and $\vf_e^{(0)} = \vf_e$.
To aid future analysis, we will represent the features as matrices, denoting the vertex embedding matrix as $\mF_V^{(k)}$ and the edge embedding matrix as $\mF_E^{(k)}$.
If a superscript is not specified then it refers to the final output of the network.

First we describe older methods of GNNs to give context, then we describe the method we use in this work.
Many older methods assume we have the adjacency matrix $\mA$ of the graph known a-priori \cite{bruna2013spectral}, and can encode graph convolutions using the eigenvectors of $\mA$ (these are known as spectral methods).
However, we do not have this luxury, as the correspondence structure changes from image set to image set, and thus we use non-spectral Graph Neural Networks.
Newer models use non-spectral methods, which often ultimately amount to transforming each node with learned weights then averaging each node's representation with its neighbors, known as a message pass \citep{kipf2017semi, defferrard2016convolutional, gama2018mimo, gama2018convolutional}.
Some works such as \cite{gama2019convolutional} use pooling operations on the vertices to make the graph smaller and thus aid computation, but as we need labels on every vertex of the original graph, we cannot use this.
Most GNNs used in these works can be expressed mathematically as:
\begin{align*}
\tilde{\vf}_v^{(k+1)} =&\; \sigma\left( b^{(k)} + \mW_0^{k} \vf_{v}^{(k)} + \sum_{h=0}^H \sum_{v' \in \mathcal{N}^{h}(v)} f_{e(v,v')} \mW_{h}^{k} \vf_{v'}^{(k)} \right) \\
\end{align*}
The weights/biases $\mW_{h}^{k}$, $b^{(k)}$ are all learned, with no learning done on the edge weights $f_{e(v,v')}$. 
Note that this is just averages over $h$-hop neighborhoods, where the weights on the edges remain static through the computation.
Given that we are trying to prune edges, we add features over edges to learn which ones to prune and which to keep such as in \cite{scarselli2009graph}.

In this work we use the method and implementation described in \cite{battaglia2018relational}.
\cite{battaglia2018relational} uses message passes between node features as well as edge features, which the model can use to prune unnecessary or erroneous edges.
Therefore there is intermediate processing on the edges before information is passed to the vertices.

@@

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth correspondences but using unsupervised methods and geometric side losses.
Thus we test against ground truth correspondence graph adjacency matrices computed from the bundle adjustment output.
Our method performs better than 35 iteration of the MatchALS \citep{zhou2015multi} method, but does not perform as well as 50 iterations.
% We do not perform as well as the Projected Gradient Descent Doubly Stochastic (PGDDS) \citep{leonardos2016distributed} but we perform significantly faster than them.
We perform better than a simple eigenvalue based method \citep{pachauri2013solving}.
Note that we perform much better in $L_1$ performance rather than $L_2$, as we optimized the network weights using an $L_1$ loss.
}
\label{tab:rome16ktab}
\end{table*}

@@

For these experiments we train with the $L_1$ norm and geometric consistency losses.
We evaluate on a test set using the ground truth adjacency matrix, which we compute from the bundle adjustment given by the Rome16K dataset.
However, we do not train with the ground truth adjacency matrix, only with a noisy version of the adjacency matrix.
We also add the geometric loss (\ref{eq:geom_cost2}) which helps improve testing performance (see figure \ref{fig:geomloss}).
We use the $L_1$ and ROC AUC metrics to measure performance.
For this method to work, we need the dimension of the embedding to be at least the number of unique points in the scene.
Picking the correct number is difficult a-priori, and is a problem with all cycle consistency based methods.
Here we use the ground truth dimension of the embedding to test both our method and the baselines.

@@

\subsubsection{Graffiti Dataset} \label{sec:graffiti}
We run our trained model on the more Graffiti Dataset from the Affine Covariant Regions dataset (formatted to be able to be input to our model properly).
The Graffiti Dataset is the most common benchmark used in feature matching algorithms (e.g. \cite{leonardos2016distributed, zhou2015multi}).
The results are shown in figure \ref{fig:errorlines} in the rightmost figure.
As the graffiti dataset is very small (only 6 views total), we were not able to train on it.
We randomly permute the intra-image order of the features to add some variance - by design the GNN outputs the same result each time, while the optimization methods have a very small amount of variance.
The transferred results of Graffiti are similar to the test error of Rome16K - smaller $L_1$ error and comparable ROC error.
This shows that the GNNs trained in Rome16K generalize similarly to the optimization based methods.
